---
title: "Hands-on Exercise 8"
date: "Last updated on `r format(Sys.time(), '%d %B, %Y')`"
format:
  html:
    code-fold: false
    code-summary: "code block"
---

# Cloropleth Mapping in R

## Loading the packages
```{r}
pacman::p_load(sf, tmap, tidyverse, skimr)
```

## Importing data  
Dataset types used for this study are:  

* ESRI Shapefile: `MP14_SUBZONE_WEB_PL`
* aspatial data file: `respopagesextod2011to2020.csv`, with fields used as unique identifiers to geocode the shapefile

::: panel-tabset
## **ESRI Shapefile**
```{r}
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")

skim(mpsz)
```
## **aspatial data**
```{r}
popdata <- read_csv("data/aspatial/respopagesextod2011to2020.csv")
skim(popdata)
```

:::

## Data Preparation

Only selected variables from year 2020 will be used for the study:  

* YOUNG: age group 0 to 4 until age groyup 20 to 24,
* ECONOMY ACTIVE: age group 25-29 until age group 60-64,
* AGED: age group 65 and above,
* TOTAL: all age group, and
* DEPENDENCY: the ratio between young and aged against economy active group

```{r}
# Preparing the attribute dataframe

popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup()%>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
mutate(`ECONOMICALLY ACTIVE` = rowSums(.[7:11])+
rowSums(.[13:15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMICALLY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
       `ECONOMICALLY ACTIVE`, `AGED`, 
       `TOTAL`, `DEPENDENCY`)
```
### Standardising uppercase of fields 
```{r}
popdata2020 <- popdata2020 %>%
  mutate(PA = toupper(PA), 
         SZ = toupper(SZ)) %>%
  filter(`ECONOMICALLY ACTIVE` > 0)
```

### Left join on identifier
`left_join()` of dplyr is used to join the geographical data and attribute table using planning subzone name e.g. SUBZONE_N and SZ as the common identifier.
```{r}
mpsz_pop2020 <- left_join(
  mpsz, 
  popdata2020,
  by = c("SUBZONE_N" = "SZ"))
```

### write rds file
```{r}
write_rds(mpsz_pop2020, "data/rds/mpszpop2020.rds")
```

## Choropleth Mapping Geospatial Data Using `tmap`

### Using `qtm()`

This creates a quick cartographic standard choropleth map:  

* `tmap_mode()` with “plot” option is used to produce a static map. For interactive mode, “view” option should be used.
* `fill` argument is used to map the attribute (i.e. DEPENDENCY)

```{r}
tmap_mode("plot")
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY")
```

##  Using tmap’s elements

`tmap` has highly customizable options, with more control over map aesthetics:
```{r}
#| fig-width: 8

tm_shape(mpsz_pop2020)+
  tm_fill(
    "DEPENDENCY", 
    style = "quantile", 
    palette = "Purples",
    title = "Dependency ratio"
  ) +
  tm_layout(
    main.title = "Distribution of Dependency Ratio by planning subzone",
    main.title.position = "center",
    main.title.size = 1.2,
    legend.height = 0.45, 
    legend.width = 0.3,
    frame = FALSE,
    bg.color = "#F8F3E6"
  ) +
  tm_borders(alpha = 0.3) +
  tm_compass(
    type="8star", 
    size = 1
  ) +
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
  tm_credits(
    "Source: Planning Sub-zone boundary from URA\n and Population data from Department of Statistics DOS", 
   position = c("left", "bottom")
  )

```
### tmap: custom breaks

Understanding distribution of **DEPENDENCY** values:

```{r}
summary(mpsz_pop2020$DEPENDENCY)
```

Adding `breaks` argument to `tm_fill()`:

```{r}
#| warning: false
tm_shape(mpsz_pop2020)+
  tm_fill(
    "DEPENDENCY",
    breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)
  ) +
  tm_layout( bg.color = "#F8F3E6") +
  tm_borders(alpha = 0.5)
```

### Assigning custom colors using `RcolorBrewer` palettes:

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          # Use '-' to reverse color
          palette = "-Purples"
  ) +
  tm_layout( bg.color = "#F8F3E6") +
  tm_borders(alpha = 0.5)
```

### Adding map legends

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "jenks", 
          palette = "Purples", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone \n(Jenks classification)",
            main.title.position = "center",
            main.title.size = 1,
            legend.height = 0.45, 
            legend.width = 0.35,
            legend.outside = FALSE,
            legend.position = c("right", "bottom"),
            frame = FALSE,
             bg.color = "#F8F3E6") +
  tm_borders(alpha = 0.5)
```

### Changing map styles:
```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```

### Adding Cartographic elements

tm_compass(), tm_scale_bar() and tm_grid() are used to add compass, scale bar and grid lines onto the choropleth map:

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "No. of persons") +
  tm_layout(main.title = "Distribution of Dependency Ratio \nby planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = FALSE,
            bg.color = "#F8F3E6" ) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

NOTE: to reset default map style, use:

```{r}
tmap_style("white")
```
## Facet Mapping

In tmap, small multiple maps can be plotted in three ways:

* by assigning multiple values to at least one of the asthetic arguments,
* by defining a group-by variable in tm_facets(), and
* by creating multiple stand-alone maps with tmap_arrange().


### Assigning multiple `fill` values:

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues", "Purples") 
  ) +
  tm_layout(legend.position = c("right", "bottom"),
            bg.color = "#F8F3E6",
            frame = FALSE) +
  tm_borders(alpha = 0.5) +
  tmap_style("white")

```

### using `facet_map()` argument

```{r}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Purples",
          thres.poly = 0) + 
  tm_facets(by="REGION_N", 
            free.coords=TRUE, 
            drop.shapes=TRUE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20,
            bg.color = "#F8F3E6") +
  tm_borders(alpha = 0.5)
```

### Creating multiple maps using `tm_arrange()`

```{r}
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Purples")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)

```

## Conditional Mapping

```{r}
tm_shape(
  # Only plotting dependency for central region
  mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ]
  )+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Purples", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE,
            bg.color = "#F8F3E6") +
  tm_borders(alpha = 0.5)

```
