---
title: "Hands-on Exercise 3"
date: "Last updated on `r format(Sys.time(), '%d %B, %Y')`"
format:
  html:
    code-fold: true
    code-summary: "code block"
---
# Load and Install R Packages
```{r, message=FALSE}
pacman::p_load(ggiraph, plotly, gganimate,
               DT, tidyverse, patchwork,
               urbnthemes, ggthemr)
```

# Importing the data
```{r, message=FALSE}
exam_data <- read_csv("data/Exam_data.csv")
```

# Interactive Data Visualisation
## ggiraph package

::: {.focusbox .focus data-latex="focus"}
**Interactive Arguments of ggiraph**

-   **Tooltip:** displayed when hovering over chart elements
-   **Onclick:** executed when elements are clicked
-   **Data_id:** Uses columns as id to show association of elements
:::

### 1: Tooltips

Original ggplot dotplot:
```{r}
set_urbn_defaults(style = "print")

ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_dotplot(stackdir = "up", 
               binwidth = 1,
               dotsize = 1) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(x = "Math Scores") + 
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```
With ggiraph tooltips:
```{r}
p <- ggplot(data = exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(tooltip = ID),
    stackgroups = TRUE, 
    binwidth = 1,
    dotsize = 1,
    method = "histodot") +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(x = "Math Scores")

girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618
)

```

### 1.1: Controlling Information in Tooltips

Customise information to be included in tooltips by creating a new list

```{r}
exam_data$tooltip <- c(paste0(
  "Name:", exam_data$ID,
  "\n Class:", exam_data$CLASS
))

p <- ggplot(data = exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(tooltip = exam_data$tooltip),
    stackgroups = TRUE, 
    binwidth = 1,
    dotsize = 1,
    method = "histodot") +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(x = "Math Scores")

girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618
)

```

### 1.2: Formatting style of Tooltips

Adding css arguments to opts_tooltip() of **ggiraph**
```{r}
tooltip_css <- "background-color:#dfdfeb; font-style:bold; color:black;"  #<<

exam_data$tooltip <- c(paste0(
  "Name:", exam_data$ID,
  "\n Class:", exam_data$CLASS
))

p <- ggplot(data = exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(tooltip = exam_data$tooltip),
    stackgroups = TRUE, 
    binwidth = 1,
    dotsize = 1,
    method = "histodot") +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(x = "Math Scores")

girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618,
  options = list(    #<<
    opts_tooltip(    #<<
      css = tooltip_css)) #<<
)                               

```
### 1.3: Displaying Statistics in tooltips

Adding stat_summary() calculations in ggplot
```{r}
ggthemr("flat")

tooltip <- function(y, ymax, accuracy = .01) {   #<<
  mean <- scales::number(y, accuracy = accuracy) #<<
  sem <- scales::number(ymax - y, accuracy = accuracy) #<<
  paste("Mean Math Score:", mean, "+/-", sem) #<<
} #<<

p2 <- ggplot(data=exam_data, 
                   aes(x = RACE)) +
  stat_summary(aes(y = MATHS, 
                   tooltip = after_stat(tooltip(y, ymax))),
               fun.data ="mean_se",
               geom = GeomInteractiveCol,
               fill = "light blue") +
  stat_summary(aes(y = MATHS),
               fun.data = mean_se,
               geom ="errorbar", 
               width = 0.2, linewidth = 0.2) +
  labs(y ="Math Scores") + 
  theme(axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.background = element_rect(fill = "#F8F3E6", color = "#F8F3E6"))

girafe(ggobj = p2,
       width_svg = 8,
       height_svg = 8*0.618)

```

### 2: Hover effect with data_id 

::: {.pinbox .pin data-latex="pin"}
**data_id** is specified as an **aes() argument** in the interactive geom functions  

e.g: geom_dotplot_interactive(aes(data_id = variablename))
:::

```{r}
p <- ggplot(data = exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(           
    aes(data_id = CLASS),             #<<
    stackgroups = TRUE,               
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL) + 
  labs(x ="Math Scores") + 
  theme(axis.line = element_blank(),
        plot.background = element_rect(fill = "#F8F3E6", color = "#F8F3E6"))

girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618                      
)        
```

### 2.1: Formatting Style of Hover Effect

::: {.focusbox .focus}
**Customisation Options**  

-   **opts_hover**(css = NULL, reactive = FALSE, nearest_distance = NULL)  > effect on geometries  
-   **opts_hover_inv**(css = NULL) > effect on other geometries, while one is hovered
-   **opts_hover_key**(css = NULL, reactive = FALSE) > interactive scales
-   **opts_hover_theme**(css = NULL, reactive = FALSE) > interactive theme elements
:::

```{r}
p <- ggplot(data = exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(           
    aes(data_id = CLASS),             
    stackgroups = TRUE,               
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL) + 
  labs(x ="Math Scores") + 
  theme(axis.line = element_blank(),
        plot.background = element_rect(fill = "#F8F3E6", color = "#F8F3E6"))

girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(                     #<<
  opts_hover(css = "fill: #202020;"), #<<
  opts_hover_inv(css = "opacity:0.2;") #<<
  )                                   #<<  
) 
```

### 1 + 2: Tooltip + Hover Powers Combined!!

```{r}
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(tooltip = CLASS, #<<
        data_id = CLASS),#<<              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL) +
  labs(x ="Math Scores") + 
  theme(axis.line = element_blank(),
        plot.background = element_rect(fill = "#F8F3E6", color = "#F8F3E6"))

girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(                        
    opts_hover(css = "fill: #202020;"),  
    opts_hover_inv(css = "opacity:0.2;") 
  )                                        
)                                        
```
