---
title: "Hands on Exercise 2"
date: "Last updated on `r format(Sys.time(), '%d %B, %Y')`"
---

# Load and Install R Packages

```{r}
pacman::p_load(ggrepel, patchwork, ggpmisc,
               ggthemes, hrbrthemes, InraeThemes, ggthemr,
               tidyverse) 
```

# Importing the data

```{r, message=FALSE}
exam_data <- read_csv("data/Exam_data.csv")
```

# Overview of the data

```{r}
summary(exam_data)
```

# Plotting Practice

## 1: Annotations with ggrepel

A comparison between annotations using geom_label(), geom_text() and package ggrepel

::: panel-tabset
### with ggrepel

```{r, warning=FALSE}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              linewidth=0.5,
              formula = y~x) +  
  geom_label_repel(aes(label = ID), 
                   fontface="bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  labs(x="Math",
       y="English",
       title="Correlation between English and Maths scores for Primary 3",
       subtitle="There is a positive correlation between English and Math Scores") +
  theme(text=element_text(family="mono"),
        plot.background=element_rect(fill="#F8F3E6",colour="#F8F3E6"))
```

### without

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              linewidth=0.5,
              formula = y~x) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  labs(x="Math",
       y="English",
       title="Correlation between English and Maths scores for Primary 3",
       subtitle="There are too many annotations to see the points clearly") +
  theme(plot.background=element_rect(fill="#F8F3E6",colour="#F8F3E6"))
```
:::

## 2: Customising themes

### 2.1: **ggtheme** package

::: panel-tabset
### Plot
```{r}
#| echo: false
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  labs(x = "Math Scores",
       y = "Number of Students",
       title = "Distribution of Math Scores") +
  theme_economist()
```

### Code

```{r}
#| fig-show: hide
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  labs(x = "Math Scores",
       y = "No. of Students",
       title = "Distribution of Math Scores") +
  theme_economist()
```
:::

### 2.2: **hbrthemes** package

::: panel-tabset
### Plot 1
```{r}
#| echo: false
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  labs(x = "Math Scores",
       y = "Number of Students",
       title = "Distribution of Math Scores") +
  theme_ipsum()
```

### Code 1
```{r}
#| fig-show: hide
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  labs(x = "Math Scores",
       y = "No. of Students",
       title = "Distribution of Math Scores") +
  theme_ipsum()
```
### Plot 2
```{r}
#| echo: false
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  labs(x = "Math Scores",
       y = "Number of Students",
       title = "Distribution of Math Scores") +
  theme_ipsum(axis_title_size = 18,
              base_size = 12,
              grid = "Y")
```

### Code 2
```{r}
#| fig-show: hide
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  labs(x = "Math Scores",
       y = "Number of Students",
       title = "Distribution of Math Scores") +
  theme_ipsum(axis_title_size = 18,
              base_size = 12,
              grid = "Y")
```

:::

## 3: Composite Graphs

### 3.1 Combining 2 graphs: '+' operator

```{r}
#| fig-width: 12
#| fig-height: 8
p1 <- ggplot(data=exam_data, 
            aes(x = MATHS)) +
        geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey90", 
                 fill="#3A3B60") +
        labs(x = "Math Scores",
             y = "Number of Students",
             title = "Distribution of Math Scores")

p2 <- ggplot(data=exam_data, 
            aes(x = ENGLISH)) +
        geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey90", 
                 fill="#4d5887") +
        labs(x = "English Scores",
             title = "Distribution of English Scores") +
        theme(axis.title.y = element_blank())

p1 + p2
```

### 3.2 Combining 3 graphs: '|', '/', '()'
  
::: {.focusbox .focus data-latex="focus"}
**Use different operators to specify composition**  

*   “|” operator to stack two ggplot2 graphs
*   “/” operator to place the plots beside each other
*   “()” operator defines the sequence of the plotting
:::

```{r}
#| fig-width: 12
#| fig-height: 8
p3 <- ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
      geom_point() +
      geom_smooth(method=lm, 
              linewidth=0.5,
              formula = y~x) + 
      stat_poly_line() +
      stat_poly_eq() +
      coord_cartesian(xlim=c(0,100),
                      ylim=c(0,100)) +
      labs(x="Math",
           y="English",
           title="Correlation between Math and English Scores")

(p1/p2) | p3

```

### 3.3: Adding Annotations

```{r}
#| fig-width: 12
#| fig-height: 8
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')
```

### Applying themes to composite 

`theme_economist()`

```{r}
#| fig-width: 12
#| fig-height: 8
patchwork <- ((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')

patchwork & theme_economist()
```
`theme_inrae()`

```{r}
#| fig-width: 12
#| fig-height: 8
#| warning: false
patchwork <- ((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')

patchwork & theme_inrae()
```
  
`theme_solarized_2()`

```{r}
#| fig-width: 12
#| fig-height: 8
#| warning: false
patchwork <- ((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')

patchwork & theme_solarized_2()
```

**ggthemr: chalk**

```{r}
#| fig-width: 12
#| fig-height: 8
#| warning: false

ggthemr("chalk")

patchwork <- ((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')

patchwork
```


### 3.4: Inset Figures  
  
Using `inset_element()` to place a plot within another plot

```{r}
ggthemr_reset()

p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```